\documentclass[a4paper,10pt]{article}
\usepackage{xcolor,geometry,amsmath,tikz}
\usepackage{everypage}

% Set page background to black and text color to white
\pagecolor{black}
\color{white}

% Reduce margins to use more space
\geometry{top=1.5cm, bottom=1.5cm, left=2cm, right=2cm}

% Load TikZ calc library
\usetikzlibrary{calc}

% Watermark configuration
\AddEverypageHook{%
  \begin{tikzpicture}[remember picture,overlay]
    \node[opacity=0.7,font=\small] at ($(current page.south east)+(-1.5cm,0.5cm)$) { Omar Abbas};
  \end{tikzpicture}%
}

\begin{document}

\section*{\centering Simple Discretization Methods: Binning}

\subsection*{1. Equal-Width (Distance) Partitioning}

\textbf{Formula:}
\[
W = \frac{B - A}{N}
\]
where:
\begin{itemize}
    \item \( W \) = Width of each interval  
    \item \( A \) = Minimum value in the dataset  
    \item \( B \) = Maximum value in the dataset  
    \item \( N \) = Number of bins  
\end{itemize}

\textbf{Example:}  
Given the dataset:  
\[
\{1, 3, 7, 10, 15, 18, 22, 25\}
\]
Let \( N = 3 \), \( A = 1 \), \( B = 25 \).

\textbf{Step 1: Compute Interval Width}
\[
W = \frac{25 - 1}{3} = \frac{24}{3} = 8
\]

\textbf{Step 2: Define Bins}
\begin{itemize}
    \item \textbf{Bin 1:} \([1, 9)\) → contains \(\{1, 3, 7\}\)
    \item \textbf{Bin 2:} \([9, 17)\) → contains \(\{10, 15\}\)
    \item \textbf{Bin 3:} \([17, 25]\) → contains \(\{18, 22, 25\}\)
\end{itemize}

This method is \textbf{simple and intuitive}, but \textbf{outliers} and \textbf{skewed data} can make it less effective.

\vspace{0.5cm}

\subsection*{2. Equal-Depth (Frequency) Partitioning}

\textbf{Concept:}  
Each bin should have \textbf{approximately the same number of elements}.

\textbf{Example:}  
Given the dataset:
\[
\{1, 3, 7, 10, 15, 18, 22, 25\}
\]
Let \( N = 3 \). Since there are \textbf{8 elements},  
each bin should have about \(\frac{8}{3} \approx 3\) elements.

\textbf{Step 1: Sort Data}  
\[
\{1, 3, 7, 10, 15, 18, 22, 25\}
\]

\textbf{Step 2: Assign Bins}
\begin{itemize}
    \item \textbf{Bin 1:} \(\{1, 3, 7\}\)
    \item \textbf{Bin 2:} \(\{10, 15, 18\}\)
    \item \textbf{Bin 3:} \(\{22, 25\}\)
\end{itemize}

This method ensures \textbf{equal distribution} of data points in each bin,  
making it \textbf{better for skewed data}.

\newpage % Start a new page

\section*{\centering Bin Processing Methods}

\subsection*{1. Bin Means}
\textbf{Concept:}  
Replace all values in a bin with the \textbf{mean} (average) of that bin.

\textbf{Example:}  
Using the same bins from equal-width partitioning:

\begin{itemize}
    \item \textbf{Bin 1:} \(\{1, 3, 7\}\) → Mean = \(\frac{1+3+7}{3} = 3.67\) → \(\{3.67, 3.67, 3.67\}\)
    \item \textbf{Bin 2:} \(\{10, 15\}\) → Mean = \(\frac{10+15}{2} = 12.5\) → \(\{12.5, 12.5\}\)
    \item \textbf{Bin 3:} \(\{18, 22, 25\}\) → Mean = \(\frac{18+22+25}{3} = 21.67\) → \(\{21.67, 21.67, 21.67\}\)
\end{itemize}

\textbf{Pros:} Reduces variance.  
\textbf{Cons:} May hide important variations in data.

\vspace{0.5cm}

\subsection*{2. Bin Medians}
\textbf{Concept:}  
Replace all values in a bin with the \textbf{median} of that bin.  
The \textbf{median} is the middle value when the numbers are sorted. If there is an even number of elements, the median is the average of the two middle values.

\textbf{Example:}
\begin{itemize}
    \item \textbf{Bin 1:} \(\{1, 3, 7\}\) → Sorted: \(\{1, 3, 7\}\), Median = \(3\) → \(\{3, 3, 3\}\)
    \item \textbf{Bin 2:} \(\{10, 15\}\) → Sorted: \(\{10, 15\}\), Median = \(\frac{10 + 15}{2} = 12.5\) → \(\{12.5, 12.5\}\)
    \item \textbf{Bin 3:} \(\{18, 22, 25\}\) → Sorted: \(\{18, 22, 25\}\), Median = \(22\) → \(\{22, 22, 22\}\)
\end{itemize}

\textbf{Pros:} Handles outliers better than means.  
\textbf{Cons:} May not capture overall distribution well.

\vspace{0.5cm}

\subsection*{3. Bin Boundaries}
\textbf{Concept:}  
Replace each value with the nearest boundary value in the bin.

\textbf{Example:}
\begin{itemize}
    \item \textbf{Bin 1:} \(\{1, 3, 7\}\) → Boundaries: \(1, 7\)  
          - \(1 \to 1\), \(3 \to 1\), \(7 \to 7\) → \(\{1, 1, 7\}\)
    \item \textbf{Bin 2:} \(\{10, 15\}\) → Boundaries: \(10, 15\)  
          - \(10 \to 10\), \(15 \to 15\) → \(\{10, 15\}\)
    \item \textbf{Bin 3:} \(\{18, 22, 25\}\) → Boundaries: \(18, 25\)  
          - \(18 \to 18\), \(22 \to 25\), \(25 \to 25\) → \(\{18, 25, 25\}\)
\end{itemize}

\textbf{Pros:} Good for preserving extreme values.  
\textbf{Cons:} Can distort middle-range values.

\newpage % Start a new page

\section*{\centering Data Transformation: Normalization}

\subsection*{1. Min-Max Normalization}

\textbf{Formula:}
\[
v' = \frac{v - \text{min}_A}{\text{max}_A - \text{min}_A} \times (\text{new\_max}_A - \text{new\_min}_A) + \text{new\_min}_A
\]
where:
\begin{itemize}
    \item \( v \) = Original value  
    \item \( \text{min}_A \) = Minimum value of the original dataset  
    \item \( \text{max}_A \) = Maximum value of the original dataset  
    \item \( \text{new\_min}_A \) = Minimum value of the new range  
    \item \( \text{new\_max}_A \) = Maximum value of the new range  
    \item \( v' \) = Normalized value
\end{itemize}

\textbf{Example:}  
Convert age = 30 to a range [0, 1], where \( \text{min} = 10 \) and \( \text{max} = 80 \):

\[
v' = \frac{30 - 10}{80 - 10} = \frac{20}{70} = \frac{2}{7} \approx 0.2857
\]

This transformation scales the value to fall between the new range, in this case, between 0 and 1.

\vspace{0.5cm}

\subsection*{2. Z-Score Normalization}

\textbf{Formula:}
\[
v' = \frac{v - \text{mean}_A}{\text{stand\_dev}_A}
\]
where:
\begin{itemize}
    \item \( v \) = Original value  
    \item \( \text{mean}_A \) = Mean of the original dataset  
    \[
    \text{mean}_A = \frac{1}{n}\sum_{i=1}^{n} x_i
    \]
    \item \( \text{stand\_dev}_A \) = Standard deviation of the original dataset  
    \[
    \text{Population: } \sigma = \sqrt{\frac{1}{n}\sum_{i=1}^{n} (x_i - \text{mean}_A)^2}
    \]
    \[
    \text{Sample: } s = \sqrt{\frac{1}{n-1}\sum_{i=1}^{n} (x_i - \text{mean}_A)^2}
    \]
    \item \( v' \) = Normalized value
\end{itemize}

The Z-score normalization transforms the data to have a \textbf{mean of 0} and a \textbf{standard deviation of 1}, making it useful for comparing data from different distributions.

\vspace{0.5cm}

\subsection*{3. Normalization by Decimal Scaling}

\textbf{Formula:}
\[
v' = \frac{v}{10^j}
\]
where \( j \) is the smallest integer such that:

\[
\text{Max}\left(|v'|\right) < 1
\]

This transformation scales the data by dividing each value by a power of 10, where the exponent \( j \) is chosen to ensure the maximum absolute value of \( v' \) is less than 1.

\vspace{0.5cm}

\end{document}
